---
title: "assigment1"
output: html_document
---
#Peer reviewed assigment 1

Importing libraries.
```{r}
library(ggplot2)
```

##Reading and processing the data

```{r}
data <- read.csv('activity.csv')
data$date <- as.Date(data$date)
```

##What is mean total number of steps taken per day?

Calculate the total steps by day and plot a histogram. Calculate the mean and median of the steps by day.

```{r}
steps_by_day <- tapply(data$steps, data$date, sum, na.rm=TRUE)
hist(steps_by_day, xlab='Number of steps per day', main='Total steps by day')
steps_mean <- mean(steps_by_day)
steps_median <- median(steps_by_day)

```

The mean is `r steps_mean` and the median is `r steps_median`.

##What is the average daily activity pattern?

```{r}
average_per_interval <- aggregate(x=list(mean_steps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
plot(average_per_interval, type='l', main='Average daily activity pattern', ylab='Average number of steps taken', xlab='Interval')
most_steps <- which.max(average_per_interval$mean_steps)
time_most_steps <- gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", average_per_interval[most_steps,'interval'])

```

Most steps were taken at `r time_most_steps`.

##Imputing missing values

```{r}
missing_values <- length(which(is.na(data$steps)))
```

There are `r missing_values` rows with missing data.

Cleaning data:

```{r}
na_positions <- which(is.na(data$steps))
mean_vec <- rep(mean(data$steps, na.rm=TRUE), times=length(na_positions))
cleaned_data <- data
cleaned_data[na_positions, 'steps'] <- mean_vec
```

Plotting a histogram of the total amount of steps each day, calculating mean and median.

```{r}
cleaned_steps_by_day <- tapply(cleaned_data$steps, cleaned_data$date, sum)
hist(cleaned_steps_by_day, xlab='Number of steps per day', main='Total steps by day')
cleaned_steps_mean <- mean(cleaned_steps_by_day)
cleaned_steps_median <- median(cleaned_steps_by_day)

```

The mean is `r cleaned_steps_mean` and the median is `r cleaned_steps_median`. These values are different from before. Having more data seems to increase the mean and median.

##Are there differences in activity patterns between weekdays and weekends?

Create a factor for the cleaned data set, indicating whether the day is a weekday or a weekend day, then plot a time series plot for each kind of day.

```{r}
cleaned_data$dateType <-  ifelse(as.POSIXlt(cleaned_data$date)$wday %in% c(0,6), 'weekend', 'weekday')
activity_per_day_type <- aggregate(steps ~ interval + dateType, data=cleaned_data, mean)
ggplot(activity_per_day_type, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```